---
title: 'Reproducible Research: Peer Assessment 1'
output: html_document
keep_md: true
---
## Loading and preprocessing the data
First, we load the data and transform **date** and **interval** into factor variables.

```{r}
activity <- read.csv("~/Downloads/activity.csv", header = T)
activity = transform(activity, date = factor(date),
                     interval = factor(interval))
```

## What is mean total number of steps taken per day?
Ignoring the missing values, we got the histogram of total number of steps taken per day as below.

```{r stepsmean1}
library(dplyr)
library(lattice)
steps_mean1 <- 
  activity %>%
    group_by(date) %>%
      summarize(day_steps = sum(steps, na.rm = T)) 
hist(steps_mean1$day_steps, main = "Histogram of total number of 
      steps taken per day", xlab = "Total number of steps")
mean1 <- as.integer(mean(steps_mean1$day_steps))
median1 <- median(steps_mean1$day_steps)
```
The mean and median of the total number of steps taken per day is `r mean1` and `r median1`, respectively.

## What is the average daily activity pattern?
The time series plot of the 5-minute interval and the average number of steps taken averaged across all days is as below.

```{r stepsmean2}
steps_mean2 <- 
  activity %>%
    group_by(interval) %>%
      summarize(int_steps = mean(steps, na.rm = T)) 
xyplot(int_steps ~ interval, data = steps_mean2, type = 'l',
       xlab = "Interval", ylab = "Number of Steps")
max <-
  steps_mean2 %>%
    filter(int_steps == max(int_steps))
```
The 5-minuter interval with the maximum number of steps is `r max`, which the former value is the interval id and the later value is the maximum number of steps.

## Imputing missing values
Here I filled in the missing values with the mean for each day. The new dataset is *activity2*. The histogram is as below.

```{r stepsmean3}
activity2 <-  
  activity %>%
    group_by(date) 
activity2 <- mutate(activity, steps = ifelse(is.na(steps),
                            as.integer(mean(steps, na.rm = T)),
                            steps))
steps_mean3 <- 
  activity2 %>%
    group_by(date) %>%
      summarize(day_steps = sum(steps)) 
hist(steps_mean3$day_steps, main = "Histogram of total number of steps
     taken per day", xlab = "Total number of steps")
mean2 <- as.integer(mean(steps_mean3$day_steps))
median2 <- median(steps_mean3$day_steps)
```
The mean and median of the total number of steps taken per day is `r mean2` and `r median2`, respectively.

## Are there differences in activity patterns between weekdays and weekends?
*days_ends* is the new factor variable with the two levels: "weekday" and "weekend". Below is the panel plot.

```{r stepsmean4}
steps_mean4 <- 
  activity2 %>%
    mutate(days_ends = weekdays(as.Date(date))) %>%
      mutate(days_ends = ifelse((days_ends == "Saturday" | 
                                   days_ends == "Sunday"),
                                  "weekend","weekday")) %>%
        group_by(interval, days_ends) %>%
          summarize(int_steps = mean(steps, na.rm = T)) 
xyplot(int_steps ~ interval | days_ends, data = steps_mean4, type = 'l',
       layout = c(1,2), xlab = "Interval", ylab = "Number of Steps")
```